---
title: "Raw Data Figures"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
```{r, echo = FALSE, message=FALSE}

library(here)
library(purrr)
library(dplyr)
library(lubridate)
library(DT)
library(ggplot2)
library(plotly)
library(leaflet)
library(tgc)
library(strings)
library(RColorBrewer)


theme_set(theme_light())

dt_options <- list(
      dom = 'ft',
      paging = FALSE,
      searching = TRUE,
      scrollY = "500px",
      pageLength = 500,
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

```

```{r, echo=FALSE, message=FALSE}

# import data
dat_raw <- list.files(here("data"), pattern = "rds", full.names = TRUE) %>%
  unlist() %>%
  map_dfr(readRDS) %>%
  filter(VARIABLE == "Temperature") %>% 
   select(-VARIABLE, -SENSOR, -DEPLOYMENT_PERIOD) %>% 
  filter(STATION != "Piper Lake", STATION != "Hourglass Lake") 

dat_final <- dat_raw 

```


# Example Stations

## Long Season (> 12 months) {.tabset}

### Ingomar

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Ingomar") %>% 
  plot_temperature_at_depth() 

```

### McNutts Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "McNutts Island") %>% 
  plot_temperature_at_depth() 

```

### Blue Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Blue Island") %>% 
  plot_temperature_at_depth() 

```

### Flat Island

S1 - 5 m & S2

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Flat Island") %>% 
  plot_temperature_at_depth() 

```

### Birchy Head

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Birchy Head") %>% 
  plot_temperature_at_depth() 

```

### Shut-In Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Shut-In Island") %>% 
  plot_temperature_at_depth() 

```

### Shad Bay

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Shad Bay") %>% 
  plot_temperature_at_depth() 

```

### Kavanagh Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Kavanagh Point") %>% 
  plot_temperature_at_depth() 

```


## Medium Season (8 months to < 12 months) {.tabset}

### Old Point Channel

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Old Point Channel") %>% 
  plot_temperature_at_depth() 

```

### Ram Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Ram Island") %>% 
  plot_temperature_at_depth() 

```

### Borgles Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Borgles Island") %>% 
  plot_temperature_at_depth() 

```

### East Gibbs Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "East Gibbs Island") %>% 
  plot_temperature_at_depth() 

```

### Beaver Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Beaver Point") %>% 
  plot_temperature_at_depth() 

```

### Wine Harbour

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Wine Harbour") %>% 
  plot_temperature_at_depth() 

```

### 1307

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "1307") %>% 
  plot_temperature_at_depth() 

```

### Meal Rock

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Meal Rock") %>% 
  plot_temperature_at_depth() 

```

### The Bull

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "The Bull") %>% 
  plot_temperature_at_depth() 

```


### The Basin

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "The Basin") %>% 
  plot_temperature_at_depth() 

```


### Northeast Branch

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Northeast Branch") %>% 
  plot_temperature_at_depth() 

```

### Seal Cove

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Seal Cove") %>% 
  plot_temperature_at_depth() 

```

### Madeline Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Madeline Point") %>% 
  plot_temperature_at_depth() 

```

### Eddy Cove

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Eddy Cove") %>% 
  plot_temperature_at_depth() 

```

### Red Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Red Point") %>% 
  plot_temperature_at_depth() 

```


## Short Season (< 8 months) {.tabset}

### Borgles Island - S2

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Borgles Island") %>% 
  plot_temperature_at_depth() 

```

### Spry Harbour

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Spry Harbour") %>% 
  plot_temperature_at_depth() 

```

### Center Bay

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Center Bay") %>% 
  plot_temperature_at_depth() 

```

### Rook Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Rook Island") %>% 
  plot_temperature_at_depth() 

```

### Moose Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Moose Point") %>% 
  plot_temperature_at_depth() 

```

### Pirate Harbour

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Pirate Harbour") %>% 
  plot_temperature_at_depth() 

```








