---
title: 'TGC Model Manuscript: Update'
author: "Danielle Dempsey"
date: "`r Sys.Date()`"
output:
  html_document: default
editor_options:
  chunk_output_type: console
---
  
```{r, echo = FALSE, message=FALSE}

knitr::opts_chunk$set(echo = FALSE, fig.width=9, message=FALSE)

library(data.table)    # export example data
library(dplyr)         # data manipulation 
library(DT)            # interactive tables
library(ggplot2)       # figures
library(here)          # relative files paths
library(leaflet)       # interactive maps
library(strings)       # convert_depth_to_ordered_factor() function
library(tgc)           # season functions
library(sf)

dt_options <- list(
      dom = 'ft',
      paging = FALSE,
      searching = TRUE,
      scrollY = "500px",
      pageLength = 500,
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

```

```{r}

# import data
dat_raw <- tibble(fread(here("data/TGC_temperature_data.csv"))) %>% 
  convert_depth_to_ordered_factor()

colour_pal <- get_colour_palette(dat_raw)

# lease shapefiles
leases <- read_sf(here("leases/geo_export_b4e7c4a6-0cc3-4654-a83f-835a2eeca93c.shp")) %>% 
  filter(county == "Halifax" | county == "Lunenburg" | county == "guysborough") %>% 
  st_transform(crs = "+proj=longlat +datum=WGS84")


```

Working title: Stocking size and growth time estimates of cultured Atlantic salmon (*Salmo salar*) under different coastal temperature regimens: guiding year-round and seasonal site selection

Co-authors (order to be discussed): Gregor Reid, Danielle Dempsey, Ryan Horricks, Leah Lewis-McCrea, Toby Balch, Andre Dumas, Jack Rensel, Roland Cusack

# Example Stations

The Thermal Growth Coefficient (TGC) model will be applied to temperature data from three of the Centre for Marine Applied Research’s (CMAR) sensor string stations.

The stations were selected based on:

  - Completeness of time series (e.g., no data gaps > 2 days)
  - Length of growing season (e.g., one "long" season ($\ge$ 12 months), one "medium" season (8 - 12 months), and one "short" season (< 8 months)
  - Temperature profile (e.g., stratification, number of depths with measurements)


## Locations

```{r}

rad <- 8
alpha <- 0.75

dat_map <- dat_raw %>% 
  select(STATION, LATITUDE, LONGITUDE) %>% 
  distinct(STATION, .keep_all = TRUE) %>% 
  mutate(
    SEASON_CLASS = case_when(
      STATION == "Rook Island" ~ "Short",
      STATION == "Madeline Point" |
        STATION == "East Gibbs Island" |
        STATION == "Beaver Point" ~ "Medium",
      STATION == "Birchy Head" ~ "Long"
    )
  )

leaflet(dat_map) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = leases,
    color = "#444444", weight = 1, 
    label = ~license_le,
    labelOptions = labelOptions(permanent = FALSE),
    opacity = 1.0, fillOpacity = 0.5,
    highlightOptions = highlightOptions(
      color = "white", weight = 2, bringToFront = TRUE
    )
  ) %>% 
  addCircleMarkers(
    data = filter(dat_map, SEASON_CLASS == "Short"),
    ~LONGITUDE, ~LATITUDE, label = ~STATION,
    fillColor = "blue", stroke = FALSE, fillOpacity = alpha, radius = rad
  ) %>% 
  addCircleMarkers(
    data = filter(dat_map, SEASON_CLASS == "Medium"),
    ~LONGITUDE, ~LATITUDE, label = ~STATION,
    fillColor = "orange", stroke = FALSE, fillOpacity = alpha, radius = rad
  ) %>% 
  addCircleMarkers(
    data = filter(dat_map, SEASON_CLASS == "Long"),
    ~LONGITUDE, ~LATITUDE, label = ~STATION,
    fillColor = "red", stroke = FALSE, fillOpacity = alpha, radius = rad
  )

```
Figure 1: Stations selected to illustrate application of the TGC model - Birchy Head (red; long season), Beaver Point (yellow; medium season), Rook Island (blue; short season). Existing leases for Lunenburg, Halifax, and Guysborough counties shown in grey.

## All Data {.tabset}

These figures show all of the temperature data for each station. The TGC model will be applied to data for a single growing season (see next section).

Figure notes:

  - Red shading: heat stress (temperature $\ge$ 18 degrees Celsius)
  - Blue shading: superchill (temperature $\le$ -0.7 degrees Celsius)
  - Dashed grey line: 4 degrees Celsius

### Birchy Head

```{r}

dat_raw %>% 
  filter(STATION == "Birchy Head") %>% 
  plot_temperature_at_depth(
    colour_palette = colour_pal,
    date_labels_format = "%Y-%b"
  ) 

```

### Beaver Point

```{r}

dat_raw %>%
  filter(STATION == "Beaver Point") %>%
  plot_temperature_at_depth(
    colour_palette = colour_pal,
    date_labels_format = "%Y-%b"
  )

```

### Rook Island

```{r}

dat_raw %>%
  filter(STATION == "Rook Island") %>%
  plot_temperature_at_depth(
    colour_palette = colour_pal,
    date_labels_format = "%Y-%b"
  )

```


## Growing Season Data {.tabset}

  - The TGC model will be applied to the growing season beginning in 2018 for each depth at each station.
  - The growing season begins when the temperature exceeds 4 degrees Celsius and does not return below 4 degrees Celsius (e.g., "4-degrees trending up").
  - The growing season ends before the first observation of superchill.
  - If there is no superchill, the season is assumed to be 540 days (~ 18 months).
  - The "Season Table" tab shows the start and end date of the season, and the season duration in months for each depth/station.

```{r, message=FALSE, warning=FALSE}

dat_seasons <- st_filter_in_growing_seasons(dat_raw)

gap_table <- dat_seasons %>% 
  check_for_data_gaps(STATION, SEASON) %>% 
  group_by(STATION, SEASON, DEPTH) %>%
  summarize(TOTAL_GAP_DAYS = sum(GAP_LENGTH_DAYS)) %>%
  ungroup()

# data to plot
dat_seasons <- dat_seasons %>% 
  left_join(gap_table, by = c("STATION", "DEPTH", "SEASON")) %>% 
  filter(TOTAL_GAP_DAYS < 2) %>% 
  convert_depth_to_ordered_factor()

# summary table
seasons_table <- dat_seasons %>% 
  group_by(STATION, SEASON, DEPTH) %>% 
  summarise(START_SEASON = min(TIMESTAMP),
            END_SEASON = max(TIMESTAMP)) %>% 
  ungroup() %>% 
  mutate(
    SEASON_DAYS = difftime(END_SEASON, START_SEASON, units = "days"),
    SEASON_DAYS = as.numeric(round(SEASON_DAYS, digits = 2)),
    SEASON_MONTHS = round(SEASON_DAYS / 30, digits = 2)
  ) %>% 
  filter(SEASON != "S3") %>% 
  select(-SEASON, - SEASON_DAYS) %>% 
  arrange(STATION,  DEPTH)

```


### Birchy Head

```{r}

dat_seasons %>%
  filter(STATION == "Birchy Head" & SEASON == "S2") %>%
  plot_temperature_at_depth(
    colour_palette = colour_pal,
    date_breaks_major = "2 month",
    date_labels_format = "%Y-%b"
  )

```

### Beaver Point

```{r}

dat_seasons %>%
  filter(STATION == "Beaver Point" & SEASON == "S2") %>%
  plot_temperature_at_depth(
    colour_palette = colour_pal,
    date_breaks_major = "1 month",
    date_labels_format = "%Y-%b"
  )

```


### Rook Island

```{r}

dat_seasons %>%
  filter(STATION == "Rook Island") %>%
  plot_temperature_at_depth(
    colour_palette = colour_pal,
    date_breaks_major = "1 month",
    date_labels_format = "%Y-%b"
  )

```

### Season Table

```{r echo=FALSE}

seasons_table %>%
  mutate(START_SEASON = format(START_SEASON),
         END_SEASON = format(END_SEASON)) %>% 
  datatable(rownames = FALSE, options = dt_options)

```



## Next Step: Apply TGC Model

Apply TGC model at each example station and depth to answer the question:

What initial stocking weight is required to grow out a salmon to market size (5 kg) under these temperature conditions?


TGC model:

$$
w_0 = (w_t^{1/3} - \frac{TGC}{1000} * degreedays)^3
$$
$$degreedays = T_{Avg} * n_{days}$$

where

$w_0$ = initial stocking weight (kg)

$w_t$ = final weight after grow-out (kg)

$TGC$ = thermal growth coefficient (remedial = 0.25, average = 0.30, elite = 0.35)

$degreedays$ = degree days experienced by fish over growing season

$T_{Avg}$ = average temperature over the growing season*

$n_{days}$ = number of days in the growing season*

*Assume no growth for 24 hours after observation $\ge$ 18 degrees Celsius (observations in this window are removed and not included in $T_{Avg}$ or $n_{days}$)

