---
title: "Season Length & Data Gaps"
output: html_document
editor_options: 
  chunk_output_type: console
---
  
```{r, echo = FALSE, message=FALSE}

library(data.table)    # export example data
library(dplyr)         # data manipulation 
library(DT)            # interactive tables
library(ggplot2)       # figures
library(here)          # relative files paths
library(leaflet)       # interactive maps
library(plotly)        # interactive plots
library(purrr)         # read in data
library(RColorBrewer)  # colour scale for season length plot
library(strings)       # convert_depth_to_ordered_factor() function
library(tgc)           # season functions


theme_set(theme_light())

dt_options <- list(
      dom = 'ft',
      paging = FALSE,
      searching = TRUE,
      scrollY = "500px",
      pageLength = 500,
      columnDefs = list(list(className = 'dt-center', targets = "_all"))
)

```

```{r, echo=FALSE, message=FALSE}

# import data
dat_raw <- list.files(here("data"), pattern = "rds", full.names = TRUE) %>%
  unlist() %>%
  map_dfr(readRDS) %>%
  filter(VARIABLE == "Temperature") %>% 
  select(-VARIABLE, -SENSOR, -DEPLOYMENT_PERIOD) %>% 
  filter(STATION != "Piper Lake", STATION != "Hourglass Lake") 

```

Explore temperature data to select example stations for the TGC paper

  - Part 1: Growing Seasons
  - Part 2: Data Gaps
  - Part 3: Growing Season Length
  - Part 4: Visualize Data Series 
  - PArt 5: Final Selection

# Station Map

```{r, echo=FALSE}

dat_raw %>% 
  distinct(STATION, .keep_all = TRUE) %>% 
  leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron)%>%
  addMarkers(~LONGITUDE, ~LATITUDE, label = ~STATION)

```

# Part 1: Growing Seasons {.tabset}

Filter `dat_raw` for STATION + DEPTH data series with full season of data using the `st_filter_in_growing_seasons()` function, with `full_season = TRUE`. This will filter out data that is not part of a growing season, and assign each remaining observation a `SEASON` label.

```{r, message=FALSE}

dat_seasons <-  dat_raw %>%
  st_filter_in_growing_seasons(full_season = TRUE,
                               trend_threshold = 4,
                               superchill_threshold = -0.7,
                               max_season = 540) 

``` 

  - `r length(unique(dat_seasons$STATION))` STATIONs have data for at least 1 full season
  - `r nrow(distinct(dat_seasons, STATION, DEPTH))` STATION + DEPTH combinations have data for at least 1 full season


# Part 2: Data Gaps 

Check `dat_seasons` for data gaps for each STATION, DEPTH, and SEASON using the `check_for_data_gaps()` function.

The function returns a table with all data gaps greater than 2 hours & prints out the number of data gaps greater than 6 hours.

Table 1 sums the gap duration for each STATION, DEPTH, and SEASON.
 
```{r}

# Table 1
gap_table <- dat_seasons %>% 
  check_for_data_gaps(STATION, SEASON, gap_length = 2, gap_warning = 6) %>% 
  group_by(STATION, SEASON, DEPTH) %>%
  summarize(TOTAL_GAP_DAYS = sum(GAP_LENGTH_DAYS)) %>%
  ungroup()

```

## Visualize Data Gaps {.tabset}

### Figure 1

```{r, echo=FALSE, fig.height=5}

cols <- get_colour_palette(gap_table)

p <- gap_table %>% 
  convert_depth_to_ordered_factor() %>% 
  arrange(desc(TOTAL_GAP_DAYS)) %>% 
  mutate(INDEX = 1:n()) %>% 
  ggplot(aes(INDEX, TOTAL_GAP_DAYS, col = DEPTH,
             text = paste(
               "Station: ", STATION, "\n",
               "Depth: ", DEPTH, "\n",
               "Season: ", SEASON, "\n",
               "Gap Days: ", TOTAL_GAP_DAYS))
  ) + 
  geom_point() +
  scale_color_manual(name = "Depth (m)", values = cols) 

ggplotly(p, tooltip = "text")

```


### Figure 2

```{r, echo = FALSE, message=FALSE, fig.height=6}

gap_table_filt <- gap_table %>% 
  filter(TOTAL_GAP_DAYS > 0) %>% 
  convert_depth_to_ordered_factor() 

cols <- get_colour_palette(gap_table_filt)

p <-  gap_table_filt %>% 
  ggplot(aes(TOTAL_GAP_DAYS, STATION, fill = DEPTH,
             text = paste(
               "Station: ", STATION, "\n",
               "Depth: ", DEPTH, "\n",
               "Gap Days: ", TOTAL_GAP_DAYS))
  ) +
  geom_col(
    position = position_dodge2(preserve = "single", reverse = TRUE, padding = 0)
  ) +
  scale_fill_manual(name = "Depth (m)", values = cols) + 
  scale_y_discrete(name = "", limits = rev) +
  theme(panel.grid.major.y =  element_blank()) 

ggplotly(p, tooltip = "text") %>% 
  config(
    modeBarButtonsToRemove = list(
      "toggleSpikelines", "autoScale2d", "select2d", "lasso2d",
      "hoverCompareCartesian", "zoom2d"
    ),
    displaylogo = FALSE,
    toImageButtonOptions = list(
      format = "png",
      filename = "Gap_days_station_depth",
      width = 1100,
      height = 1100
    )) %>%
  layout(
    margin = list(b = 90, l = 90) # to fully display the x and y axis labels
  )

```

### Table 1

```{r, echo=FALSE}

gap_table %>% 
  arrange(desc(TOTAL_GAP_DAYS)) %>% 
  datatable(rownames = TRUE, options = dt_options)

```

Filter out STATION + DEPTH + SEASON with data gap greater than 2 days.

```{r}

dat_seasons <- dat_seasons %>% 
  left_join(gap_table, by = c("STATION", "DEPTH", "SEASON")) %>% 
  filter(TOTAL_GAP_DAYS < 2)

```


# Part 3: Growing Season Length

```{r, message=FALSE, echo=TRUE}

seasons_table <- dat_seasons %>% 
  group_by(STATION, SEASON, DEPTH) %>% 
  summarise(START_SEASON = min(TIMESTAMP),
            END_SEASON = max(TIMESTAMP)) %>% 
  ungroup() %>% 
  mutate(
    SEASON_DAYS = difftime(END_SEASON, START_SEASON, units = "days"),
    SEASON_DAYS = as.numeric(round(SEASON_DAYS, digits = 2)),
    SEASON_MONTHS = round(SEASON_DAYS / 30, digits = 2)
  ) %>% 
  convert_depth_to_ordered_factor() %>% 
  arrange(STATION, SEASON, DEPTH)

# can't use identify_growing_seasons() on data that is already filtered for seasons:
## for full_seasons = TRUE: returns empty dataframe because data always starts above trend_theshold and never cross superchill_threshold
## for full_seasons = FALSE, it adds additional seasons; e.g., S1 is from June 2019 to April 2020; it will add a S2 from January 1, 2020 to April 2020

```

## Visualize Growing Season Length {.tabset}

### Figure 3

```{r, echo=FALSE, message=FALSE, fig.height=8}

cols <- get_colour_palette(seasons_table)

p <- seasons_table %>%
  group_by(STATION, SEASON) %>%
  summarise(MAX_SEASON_MONTHS = max(SEASON_MONTHS)) %>%
  ggplot(aes(x = MAX_SEASON_MONTHS, y = STATION, fill = SEASON)) +
  geom_col(
    position = position_dodge2(preserve = "single", reverse = TRUE, padding = 0)
  ) +
  scale_y_discrete(name = "", limits = rev) +
  theme(panel.grid.major.y =  element_blank())

ggplotly(p)

```

### Figure 4

```{r, echo=FALSE, fig.height=12}

p <- seasons_table %>%
  ggplot(aes(x = SEASON_MONTHS, y = STATION, fill = DEPTH)) +
  geom_col(
    position = position_dodge2(preserve = "single", reverse = TRUE, padding = 0)
  ) +
  scale_fill_manual(name = "Depth (m)", values = cols) +
  scale_y_discrete(name = "", limits = rev) +
  facet_wrap(~SEASON, ncol = 1, scales = "free_y") +
  theme(panel.grid.major.y =  element_blank())

ggplotly(p)
  
```

### Table 3
```{r, echo=FALSE}

seasons_table %>%
  mutate(START_SEASON = format(START_SEASON),
         END_SEASON = format(END_SEASON)) %>% 
  arrange(STATION, SEASON, DEPTH) %>% 
  select(-SEASON_DAYS) %>% 
  datatable(rownames = FALSE, options = dt_options)

```


## Season Statistics {.tabset}

```{r}

season_summary <- seasons_table %>%
  group_by(STATION) %>%
  summarise(AVERAGE_SEASON_MONTHS = mean(SEASON_MONTHS),
            NUMBER_SEASONS = n()) %>%
  ungroup() %>%
  mutate(AVERAGE_SEASON_MONTHS = round(AVERAGE_SEASON_MONTHS, digits = 2))

mean_season <- mean(season_summary$AVERAGE_SEASON_MONTHS)

```

Mean season length is `r round(mean_season, digits = 2)` months long. 

### Figure 5

```{r, echo=FALSE}

p <- season_summary %>%
  ggplot(aes(x = AVERAGE_SEASON_MONTHS,
             y = reorder(STATION, -NUMBER_SEASONS),
             fill = factor(NUMBER_SEASONS),
             text = paste("Average Season: ", AVERAGE_SEASON_MONTHS, "\n",
                          "Number Seasons: ", NUMBER_SEASONS))) +
  geom_col(
    position = position_dodge2(preserve = "single", reverse = TRUE, padding = 0)
  ) +
  geom_vline(xintercept = mean_season) +
  scale_y_discrete(name = "", limits = rev) +
  scale_fill_manual(
    name = "Number Seasons",
    values = brewer.pal(
      n = length(unique(season_summary$NUMBER_SEASONS)), name = "Dark2"
      )
  ) +
  theme(panel.grid.major.y =  element_blank())


ggplotly(p, tooltip = "text")

```


### Table 4

```{r echo=FALSE}

season_summary %>%
  arrange(AVERAGE_SEASON_MONTHS) %>% 
  datatable(rownames = FALSE, options = dt_options)

```


# Part 4: Visualize Data Series

Based on the above analysis:

   - Long season = > 12 months (red)
   - Medium season: 8  - < 12 months (orange)
   - Short season: < 8 months (blue)

NOTE: Figures do NOT show all data for each station - only data for FULL seasons.

```{r, echo=FALSE}

seasons_table <- seasons_table %>% 
  mutate(
    SEASON_CLASS = case_when(
      SEASON_MONTHS < 8 ~ "Short",
      SEASON_MONTHS >= 8 & SEASON_MONTHS < 12 ~ "Medium",
      SEASON_MONTHS >= 12 ~ "Long"
    ),
    SEASON_CLASS = ordered(factor(SEASON_CLASS), 
                           levels = c("Short", "Medium", "Long"))
  ) %>% 
  select(STATION, SEASON, DEPTH, SEASON_MONTHS, SEASON_CLASS)

# add season length column
dat_final <- dat_seasons %>% 
  left_join(seasons_table, by = c("STATION", "SEASON", "DEPTH")) 

```


```{r, echo=FALSE}

rad <- 8
alpha = 0.5

dat_map <- dat_final %>% 
  select(WATERBODY, STATION, SEASON, SEASON_CLASS, LATITUDE, LONGITUDE) %>% 
  distinct(STATION, SEASON, .keep_all = TRUE) 

leaflet(dat_map) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(
    data = filter(dat_map, SEASON_CLASS == "Short"),
    ~LONGITUDE, ~LATITUDE, label = ~STATION,
    fillColor = "blue",
    stroke = FALSE,
    fillOpacity = alpha,
    radius = rad
  ) %>% 
  addCircleMarkers(
    data = filter(dat_map, SEASON_CLASS == "Medium"),
    ~LONGITUDE, ~LATITUDE, label = ~STATION,
    fillColor = "orange",
    stroke = FALSE,
    fillOpacity = alpha,
     radius = rad
  ) %>% 
  addCircleMarkers(
    data = filter(dat_map, SEASON_CLASS == "Long"),
    ~LONGITUDE, ~LATITUDE, label = ~STATION,
    fillColor = "red",
    stroke = FALSE,
    fillOpacity = alpha,
     radius = rad
  )
  
```


## Long Season (> 12 months) {.tabset}

### Ingomar

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Ingomar") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### McNutts Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "McNutts Island") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Blue Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Blue Island") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Flat Island 

S1 - 5 m & S2

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Flat Island") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Birchy Head

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Birchy Head") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Shut-In Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Shut-In Island") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Shad Bay

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Shad Bay") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Kavanagh Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Kavanagh Point") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```


## Medium Season (8 months to < 12 months) {.tabset}

### Old Point Channel

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Old Point Channel") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Ram Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Ram Island") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Borgles Island - S1

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Borgles Island") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### East Gibbs Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "East Gibbs Island") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Beaver Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Beaver Point") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Wine Harbour

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Wine Harbour") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### 1307

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "1307") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Meal Rock

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Meal Rock") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### The Bull

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "The Bull") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```


### The Basin

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "The Basin") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```


### Northeast Branch

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Northeast Branch") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Seal Cove

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Seal Cove") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Madeline Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Madeline Point") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Eddy Cove

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Eddy Cove") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Red Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Red Point") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```


## Short Season (< 8 months) {.tabset}

NOTE: Many "short seasons" are at deep depths; the shallower depths were filtered out because missing START_SEASON (i.e., data series begins above the 4-degree threshold)

### Borgles Island - S2

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Borgles Island") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Spry Harbour

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Spry Harbour") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Center Bay

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Center Bay") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Rook Island

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Rook Island") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Moose Point

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Moose Point") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```

### Pirate Harbour

```{r, echo=FALSE, message=FALSE}

dat_final %>% 
  filter(STATION == "Pirate Harbour") %>% 
  plot_temperature_at_depth(facet_var = "SEASON") 

```


# Part 5: Final Selection

Long season: Birchy Head

  - 6 depths 
  - No superchill at any depth for any season
  - Good example of hedge-hogging 
  
Medium Season: Madeline Point
  
  - 5 depths
  - High variability but limited stratification
  - No obvious hurricane signals (e.g., East Gibbs Island)
  
Rook Island

 - 4 depths
 - Relatively shallow compared to other "short" season examples
 - 2 m and 5 m might be "medium" seasons, but we don't have data from earlier in the season so they were filtered out

```{r, eval=FALSE}

dat_example <- dat_raw %>% 
  filter(STATION == "Birchy Head" |
           STATION == "Beaver Point" |
           STATION == "Rook Island") 

fwrite(dat_example, file = here("data/TGC_temperature_data.csv"))

#
# dat_example2 <- dat_raw %>% 
#   filter(STATION == "Beaver Point" |
#            STATION == "East Gibbs Island") 
# 
# fwrite(dat_example2, file = here("data/TGC_temperature_data2.csv"))

```




